---
title: "Stats 112 Homework 2"
author: "Owen Lin"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

```{r}
library(nlme)
library(lme4)
```


## Problem 1
```{r}
setwd("D:\\Coding\\Stats112\\Data")
NCGS = read.table("cholesterol-data.txt", na.strings=".")
# remove observations with na entries
NCGS = NCGS[complete.cases(NCGS), ]
# Give col names
names(NCGS) = c("Trt", "ID", "M0", "M6", "M12", "M20", "M24")
# label Treatment
NCGS$Trt = factor(NCGS$Trt, levels=c(2,1), labels=c("Placebo","HighDose"))
# convert to long form
NCGS.long = reshape(NCGS, varying=list(3:7), idvar="ID", timevar="Month", times=c(0,6,12,20,24), v.names="Chol", direction="long")
# change month to time 1 2 3 4
NCGS.long$Time = as.numeric(factor(NCGS.long$Month))
```

1a:  
For both High Dose and Placebo group, the mean cholesterol level generally goes up as time progresses.
```{r}
means = tapply(NCGS.long$Chol, list(NCGS.long$Month, NCGS.long$Trt), mean)
times = c(0,6,12,20,24)
plot(times, means[,1], type="l", xlab="Month", 
	ylab="Mean Serum Cholesterol (mg/dL)",
	ylim=c(220,280), main="Mean Serum Cholesterol Trajectories by Treatment",
	col="blue", lty=1, pch=15)
points(times, means[,2], type="l",
	col="red", lty=2, pch=16)
legend("topleft",c("Placebo","High-Dose"),
	col=c("blue","red"), lty=c(1,2), pch=c(15,16))
```

1b:  
$Y_{ij} = β_1 + β_2* Trt_{ij} +β_3* I(Month = 6)_{ij} +β_4* I(Month = 12)_{ij} +β_5* I(Month = 20)_{ij} +β_6* I(Month = 24)_{ij}+β_7* TrtI(Month = 6)_{ij}+β_8* TrtI(Month = 12)_{ij}+β_9* TrtI(Month = 20)_{ij} + β_{10}*TrtI(Month = 24)_{ij} + \epsilon_i$  
where $Trt_{ij} = 1$ for HighDose and = 0 for Placebo group
```{r}
mod.unst = gls(Chol ~ Trt+factor(Month)+Trt*factor(Month), data=NCGS.long, weight=varIdent(form = ~ 1 | Time),corr=corSymm(, form = ~ Time | ID))
summary(mod.unst)
```

```{r, eval = FALSE}
ttt = cbind(NCGS.long, rep(1, nrow(NCGS.long)))
head(ttt)
colnames(ttt)[6] = "newID"
mod.cons = gls(Chol ~ Trt+factor(Month)+Trt*factor(Month), data=ttt, weight=varIdent(form = ~ 1 | Time),corr=corAR1(, form = ~ Time))
summary(mod.cons)
```


1c:  
The estimated correlation between time 1 and time 2 is 0.764, whereas the estimated correlation between time 1 and time 5 is 0.606.

1d:  
```{r}
mod.ar = gls(Chol ~ Trt+factor(Month)+Trt*factor(Month), data=NCGS.long,corr=corAR1(, form = ~ Time | ID))
summary(mod.ar)
```

1e:  
The estimated correlation between time 1 and time 2 is 0.755, whereas the estimated correlation between time 1 and time 5 is $0.755^4 = 0.325$.

1f:  
I would go with the model in part b (the unstructured model) because it has a lower AIC.

1g:  
H_0: ar(1) model is better.  
H_a: unstructured model is better.  
p-value: <0.0001  
Conclusion:  We reject null and conclude that unstructured model is better, which does agree with the conclusion in part 1f.
```{r}
anova(mod.ar, mod.unst)
```

1h:  
```{r}
mod.h = gls(Chol ~ Trt + as.numeric(Month) + Trt:as.numeric(Month), data=NCGS.long, weight=varIdent(form = ~ 1 | Time),corr=corSymm(, form = ~ Time | ID), method = "ML")
summary(mod.h)
```

1i:  
$Y_{ij} = β_0 + β_1* Trt_{ij} + β_2* t_{ij} + β_3* t^{2}_{ij} + β_4* Trt_{ij}t_{ij} + β_5* Trt_{ij}t^{2}_{ij} + \epsilon_i$
```{r}
Month2 = as.numeric(NCGS.long$Month)^2
mod.exp = gls(Chol ~ Trt + as.numeric(Month) + Trt*as.numeric(Month) + Month2 + Trt*Month2, data=NCGS.long, weight=varIdent(form = ~ 1 | Time),corr=corSymm(, form = ~ Time | ID), method="ML")
summary(mod.exp)
```

1j:  
The model in part i (with quadratic terms) fit the data better based on likelihood ratio test at 95% confident level.
```{r}
anova(mod.h, mod.exp)
```

## Problem 2
2a:  
The marginal mean (population expectation) cholesterol level for a male subject on day 20 is  
$\beta_0 + \beta_1* 20 + \beta_2* 1$

2b:  
The conditional mean (conditional expectation) cholesterol level for a male subject on day 20 is
$\beta_0 + \beta_1* 20 + \beta_2* 1 + b_{0,i} + b_{1,i}* 20$

2c:  
The difference in marginal mean cholesterol levels between a male subject on day 20 and a female subject on day 20 is  
$\beta_2* 1$

2d:  
The difference in conditional mean cholesterol levels between a male subject on day 20 and a female subject on day 20 is  
$\beta_2* 1 + b_{0,i1} + b_{1,i1}* 20 - b_{0,i2} - b_{1,i2}* 20$  
where i1 represent a male subject and i2 represent a female subject

2e:  
For different individual, they could start with a different level of cholesterol, this is captured in random intercept ($b_{0,i}$). Similarly, the change in cholesterol level varies for different individual, this is captured in random slope ($b_{1,i}$).

## Problem 3
3a:  
$E(\bar{Y_i}|X) = \beta_0 + \beta_1* \bar{x_{1i}} + \beta_2* \bar{x_{2i}}$

3b:  
We should weight the variance of each instructor's average rating by the number of raters because $var(\bar{Y_i}) = \frac{1}{n_i}* var(Y_{ij})$

```{r}
library(alr4)
data(Rateprof)
gls = gls(quality ~ easiness + helpfulness, data = Rateprof)
coef(gls)
```