---
title: "Stats 112 Homework 3"
author: "Owen Lin"
output: html_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = "D:\\Coding\\Stats112\\Data")
library(nnet)
library(lattice)
library(nlme)
library(lme4)
library(survival)
```



## Problem 1
1a:  
```{r}
birthwt = read.table("birthwt-data.txt", header=FALSE)
names(birthwt) = c("MID", "Order", "Wt", "Age", "CID")
mod = lme(Wt ~ 1 + Age, random = ~ 1| MID, data = birthwt, method = "ML")
summary(mod)
```

1b:  
$y_{ij} = b_{0i} + 2785.1495 + 17.1431* Age_{ij} + \epsilon_{ij}$  
The estimate of $\sigma_{\epsilon}$ is 434.1899.

1c:  
The p-value is smaller than 0.05, so Age is a significant covariate of Wt and it should not be dropped from the model.
```{r}
mod0 = lme(Wt ~ 1, random = ~ 1 | MID, data = birthwt, method = "ML")
anova(mod0, mod)
```

1d:  
```{r}
mod1 = lme(Wt ~ 1 + Age, random = ~ 1 + Age | MID, data = birthwt, method = "ML")
summary(mod1) 
```

1e:  
$y_80 = -126.0165 + 2805.7881 + 15.9444* Age_{80, j} + 21.6985* Age_{80, j}$

1f:  
The random slope is needed on top of a random intercept as the p-value is less than 0.0001.
```{r}
mod1.reml = lme(Wt ~ 1 + Age, random = ~1+Age | MID, data=birthwt, method="REML")
mod.reml = lme(Wt ~ 1 + Age, random = ~1 | MID, data=birthwt, method="REML")
anova(mod.reml, mod1.reml)
```

## Problem 2
Within a high school, we would expect most student to have a comparable socioeconomical background, and sex is something that was initially determined at birth. Thus, it is reasonable that the variance of Math Achievement score is constant across different students at the same high school. Also, if students are picked randomly, there is no reason to say student1 and student2 has a different correlation than student 1 and student 4. Then the corrlelation between any pair of score would be the same. That is the precise definition of a compound symmetry structure.


## Problem 3
3a:  
```{r}
NCGS = read.table("cholesterol-data.txt", na.strings=".")
names(NCGS) = c("Trt", "ID", "M0", "M6", "M12", "M20", "M24")
NCGS$Trt = factor(NCGS$Trt, levels=c(2,1), labels=c("Placebo","HighDose"))
NCGS.long = reshape(NCGS, varying=list(3:7), idvar="ID", timevar="Month", times=c(0,6,12,20,24), v.names="Chol", direction="long")
mod.lme  = lme(Chol~Trt+I(Month)+I(Month^2), random=~I(Month)+I(Month^2)|ID, data=NCGS.long, na.action=na.omit, method="ML") # 1+, the intercept is default
summary(mod.lme)
```

3b:  
Treatment main effect is not significant with a high p-value of 0.8949.

3c:  
```{r}
getVarCov(mod.lme, type = "random.effects")
```

3d:  
The estimated variance of the random intercepts is $1670.5$.  
The estimated variances of the random slopes are $1.2282$ and $0.00246$.  
The estimated correlation between the random intercepts and slopes are $\frac{-30.77200}{\sqrt{1670.5* 1.228}}$ and $\frac{0.86151}{\sqrt{1670.5* 0.00246}}$

3e:  
The random effects for the subject with id = 96 are [-48.069, 0.625, -0.0229]
```{r}
random.effects(mod.lme)[96,]
```

3f:  
The estimated response for Month 6 for the subject with ID = 71 is 215.6939.
```{r}
fitted(mod.lme)[names(fitted(mod.lme))=="71"]
```

3g:  
At a 5% significance level, we should not have random effects on the time covariates with a p-value of 0.2121 from the likelihood ratio test.
```{r}
mod.lme.reml = lme(Chol~Trt+I(Month)+I(Month^2), random=~I(Month)+I(Month^2)|ID, data=NCGS.long, na.action=na.omit, method="REML")
mod.lme.notime = lme(Chol~Trt+I(Month)+I(Month^2), random=~1|ID, data=NCGS.long, na.action=na.omit, method="REML")
anova(mod.lme.notime, mod.lme.reml)
```