---
title: "Stats 111 Homework 2"
author: "Owen Lin"
output: html_document
---

```{r, echo = FALSE}
library(epitools)
library(rmeta)
library(pROC)
prop.comp <- function( x, estimate="all", conf.level=.95, transpose=FALSE ){
	if( transpose ) x <- t(x)
	rslt <- vector( "list", length=3 )
	names( rslt ) <- c( "riskdiff", "riskratio", "oddsratio" )
	diff.rslt <- suppressWarnings(prop.test( x, conf.level=conf.level ))
	rslt[[1]] <- rslt[[2]] <- rslt[[3]] <- epitab( x, method="riskratio", pvalue="chi2", conf.level=conf.level )$tab
	colnames( rslt[[1]] )[5] <- "riskdiff"
	rslt[[1]][,5] <- c(0,diff(rev(diff.rslt$estimate)))
	rslt[[1]][2,6:7] <- diff.rslt$conf.int
	colnames( rslt[[3]] )[5] <- "oddsratio"
	rslt[[3]][,5:8] <- suppressWarnings(epitab( x, method="oddsratio", pvalue="chi2", conf.level=conf.level )$tab[,5:8])
	if(is.null(names(dimnames(x)))){
		for(i in 1:3){
			colnames(rslt[[i]])[c(1,3)] <- c("Outcome=0", "Outcome=1")
			rownames(rslt[[i]]) <- c("Group=1", "Group=2")
			}
	}
	if( is.element( estimate, c("all", "oddsratio") ) ){ 
		if(is.null(names(dimnames(x)))){
			warning( "Estimated probabilities represent Pr[ Outcome | Group ]. For estimates of 
			Pr[ Group | Outcome ], change the value of 'transpose'.")
		}
		else
			warning( paste("Estimated probabilities represent Pr[", names(dimnames(x))[2], 
			"|",names(dimnames(x))[1], "]. For estimates of 
			Pr[", names(dimnames(x))[1], "|",names(dimnames(x))[2], "], change the value of 'transpose'.") )
		}
	if( estimate == "riskdiff" ) return(rslt[[1]])
	else if( estimate == "riskratio" ) return(rslt[[2]])
	else if( estimate == "oddsratio" ) return(rslt[[3]])
	else return(rslt)
}
```

1. a. $H_0:$ Cholesterol level and CHD are independent <br>
    $H_a:$ Cholesterol level and CHD are dependent
b. test statistic: 35.028 ~ chisquared(3) <br>
    p-value: $1.202*10^{-7}$
```{r, results = 'hold'}
heart = matrix(c(307, 246, 439, 245, 12, 8, 31, 41), 4, 2)
rownames(heart) = c("low", "above normal", "high", "very high")
colnames(heart) = c("CHD no", "CHD yes")

# chi square test of independence
chisq.test(heart , correct=FALSE)
```

c. conclusion: Cholesterol level and CHD are dependent at 5% significant level
d. It is natural to think that higher cholesterol tends to lead to heart disease. So "low", "above normal", "high", "very high" comes in a natural order.
e. $H_0:$ Probability of CHD didn't increase/decrease with Cholesterol level. <br>
    $H_a:$ H0 is not true
f. test statistic: 26.167 ~ chisquared(1)<br>
    p-value: $3.131*10^{-7}$<br>
    conclusion: We reject the null and conclude that there is a trend between CHD and Cholesterol level.
```{r}
# addmargins(heart)
prop.trend.test(heart[ ,2], rowSums(heart))
heart[ , 2]/rowSums(heart)
```

2. a. I use chi-squared of independence and got a p-value < $2.2*10^{-16}$. So there is evidence that socioeconomic level is correlated with delinquent status.
```{r, results = 'hold'}
# This is the 3 way scouts table
scout = array( c(169,43,42,11,59,196,10,10),dim = c(2, 2, 2),dimnames = list( Scout = c("No","Yes"),Verdict = c("Delinquet No", "Delinquet Yes"),Socioeconomic = c("Lomessw","High")))

# This is the two way table of Socioeconomc status and Scout status
socio_scout = array(c(211,69,54,206), dim=c(2,2),dimnames=list(Socioeconomic=c("Low","High"), Scout=c("No","Yes")))

chisq.test(socio_scout, correct=FALSE)
```

b. I use odds ratio estimate and got a odds ratio of 0.314. So there is evidence that socioeconomic level is correlated with scout status.
```{r}
# This is the two way table of Socioeconomic stats and Deliquency status
socio_deliq = array(c(212,255,53,20), dim=c(2,2),dimnames=list(Socioeconomic=c("Low","High"), Deliquent=c("No","Yes")))

chisq.test(socio_deliq, correct = F)
prop.comp(socio_deliq)
```

c. Socioeconomic status looks like a confounder in the association between Scout status and Delinquency status. Because it is related to both Scout and Deliquency Status.
d. OR_MH = 0.657, it's rather away from 1, so it seems that delinquet and scout status are dependent when controling for socioeconomic level.
# Adjusting for socioeconomic status, the odds ratio for comparing scout yes to scout no is 0.657. Adjusting for socioeconomic status, the odds of being delinquent for the scout yes is 0.657 times that of the odds of being delinquent for the scout no group.
```{r}
OR_MH = (169*11/(169+42+43+11)+59*10/(59+10+196+10))/(43*42/(169+42+43+11)+196*10/(59+10+196+10)) 
```
e. H0: OR_MH = 1
    Ha: OR_MH != 1
    p-value: 0.1843
    conclusion: we fail to reject the null => ie. delinquet and scout status are uncorrelated when controling for socioeconomic levels.
```{r}
mantelhaen.test(scout)
```
# Run BD Test to check the assumption from MH test that the odds ratios with each stratum of W are approximately equal (Thus we can combine them to create a single common odds ratio)
```{r}
# two way table ignoring Socioeconomic status
scout_margin = scout[,,1]+ scout[,,2]

# Compute relevant counts to conduct odds ratio heterogeneity test
n.scout = c(sum(scout[,,1][2,]), sum(scout[,,2][2,]))
n.no = c(sum(scout[,,1][1,]), sum(scout[,,2][1,]))
s.deliq = c(scout[,,1][2, 2], scout[,,2][2, 2])
n.deliq= c(scout[,,1][1, 2], scout[,,2][1, 2])

# # Conduct Maentel Haenszel test
mantelhaen.test(scout)

# # Test for heterogeneity with the Breslow-Day test
mh.rslt = meta.MH(n.scout, n.no, s.deliq, n.deliq, names = c("Socio low", "Socio High"))
summary( mh.rslt )
# BD with pvalue = 0.04 => reject odds ratio is the same across diff W => shouldn't do MH test
```

f. The effect of being delinquent on scout status differs across different socioeconomic levels. So socioeconomic level is a effect modifier.

3. a. If the person doesn't drink and smoke, then the expected BP is beta_0. It is a meaningful intepretaion because it is a plausible condition.
b. If the person Smoke, then drink brings additional beta_3 effect on BP.
c. The effect of drinking on BP for smokers is beta2_hat + beta3_hat more than nonsmokers.
d. H0: beta2 = beta3 = 0 <br>
    Ha: H0 is not true <br>
    Reduced Model: y = beta0 + beta1xDrink <br>
    Full Model: y = beta0 + beta1xDrink + beta2xSmoke + beta3xDrinkxSmoke

4. a. BMI reasonably correlates with blood pressure and the prevalence of CHD.
b. Both p-values are way below 0.05 significant level. That means BMI is a confounder. <br>
H0: SBP and BMI are independent. <br>
Ha: SBP and BMI are dependent. <br>
H0: CHD and BMI are independent. <br>
Ha: CHD and BMI are dependent.
```{r, results = "hold"}
# Framingham Data
framingham = read.table("C:\\Users\\Owen\\Downloads\\Framingham copy.txt")
#####	Recode sex to something obvious (sex=1 -> male)
framingham$sex = framingham$sex - 1
names( framingham )[1] = "female"

#####	Create SBP and BMI groups
framingham$sbphi = cut( framingham$sbp, breaks=c(min(framingham$sbp),146, max(framingham$sbp)), include.lowest=TRUE )
framingham$bmigrp = cut( framingham$bmi, breaks=c(min(framingham$bmi, na.rm=TRUE),20, 25, 30, max(framingham$bmi, na.rm=TRUE)), include.lowest=TRUE, right=FALSE )

bmisbp.table = xtabs( ~ bmigrp + sbphi, data=framingham )
epitab( bmisbp.table, pvalue="chi2" )
n.hisbp = bmisbp.table[,2]
n.strata = rowSums(bmisbp.table) 
chisq.test(bmisbp.table)
prop.trend.test( n.hisbp, n.strata )

#####	Compute test of independence and test for trend for BMI and CHD
bmichd.table = xtabs( ~ bmigrp + chdfate, data=framingham )
epitab( bmichd.table, pvalue="chi2" )
n.chd = bmichd.table[,2]
n.strata = rowSums(bmichd.table)
chisq.test(bmichd.table)
prop.trend.test(n.chd, n.strata)
```

c. We can see that the odds of CHD are significantly greater(OR>1) for individuals with a high systolic BP, regardless of BMI group.
```{r}
#####	Compute relevant counts
n.sbplo = xtabs( ~ sbphi + bmigrp, data=framingham )[1,]
n.sbphi = xtabs( ~ sbphi + bmigrp, data=framingham )[2,]
lo.chd = xtabs( ~ chdfate + sbphi + bmigrp, data=framingham )[2,1,]
hi.chd = xtabs( ~ chdfate + sbphi + bmigrp, data=framingham )[2,2,]

# Just to show the 3 way table that is being used above for the MH and BD tests
sbphi.chd.table = table(BP=framingham$sbphi, CHD=framingham$chdfate, BMI=framingham$bmigrp)

#####	Compute M-H estimate of adjusted OR and test for heterogeneity
mh.rslt = meta.MH(n.sbphi, n.sbplo, hi.chd, lo.chd, names=levels(framingham$bmigrp))
summary( mh.rslt )
```

d. The MH common odds ratio is 1.66 with a 95% CI of (1.43, 1.92). Because 1 is not in the CI, so CHD and sbp are related while adjusting for the level of BMI.
```{r}
#####	Compute relevant counts
n.sbplo = xtabs( ~ sbphi + bmigrp, data=framingham )[1,]
n.sbphi = xtabs( ~ sbphi + bmigrp, data=framingham )[2,]
lo.chd = xtabs( ~ chdfate + sbphi + bmigrp, data=framingham )[2,1,]
hi.chd = xtabs( ~ chdfate + sbphi + bmigrp, data=framingham )[2,2,]

# Just to show the 3 way table that is being used above for the MH and BD tests
sbphi.chd.table = table(BP=framingham$sbphi, CHD=framingham$chdfate, BMI=framingham$bmigrp)

#####	Compute M-H estimate of adjusted OR and test for heterogeneity
mh.rslt = meta.MH(n.sbphi, n.sbplo, hi.chd, lo.chd, names=levels(framingham$bmigrp))
summary(mh.rslt)
# Test for heterogeneity: X^2( 3 ) = 4.03 ( p-value 0.2579 ). Fail to reject null => OR are the same across W.

# MH test using 3 way table
mantelhaen.test(sbphi.chd.table)
```

e.  H0: $OR_{MH} = 1$ <br>
    Ha: $OR_{MH} \neq 1$ <br>
    p-value: $1.268*10^{-11}$ <br>
    conclusion: We reject the null and conclude that CHD and sbp are dependent while adjusting for BMI.